The Chemistry Development Kit (CDK) \cdkversion{}: atom typing, depiction, molecular formulas, and substructure searching

### 2017-05-07 article.tex
Clarified what the Blue Obelisk is (to improve the impact of the paper outside the journal; closes #79)

The open source cheminformatics community has made significant steps
forward recently~\cite{OBoyle2011b} as evidenced by the growing number
of tools and underlying toolkits, along with the usage of these
software components in a variety of applications.
The Chemistry Development Kit (CDK) is one of the tools developed
under the aegis of the Blue Obelisk 
and previously documented versions have been widely adopted~\cite{Steinbeck2003,Steinbeck2006}.
Use of the CDK ranges from inclusion of CDK functionality in
wrapper platforms such as Cinfony~\cite{OBoyle2008}, incorporation
within the R environment (rcdk~\cite{Guha2007}), and as plugins for
Taverna~\cite{Truszkowski2011}, 
KNIME~\cite{Beisken2013}, Cytoscape (ChemViz2~\cite{ChemViz2}), and for
Microsoft Excel (LICSS~\cite{Lawson2012}).
In contrast to scenarios that have made CDK functionality available in
larger systems, a number of projects have employed the CDK as a
general cheminformatics toolkit. Examples include 
jCompoundMapper~\cite{Hinselmann2011}, ScaffoldHunter~\cite{wetzel2009interactive,Klein2013}, OMG~\cite{Peironcely2012},
PaDEL~\cite{yap2011padel}, ChemDes~\cite{Dong2015},
ReactPRED~\cite{ReactPRED}, SMSD~\cite{Rahman2009,Rahman2014,Rahman2016},
WhichCyp~\cite{Rostkowski2013}, MetaPrint2D~\cite{Carlsson2010}, MetFrag~\cite{Wolf2010},
and the IUPHAR/BPS Guide to Pharmacology~\cite{Southan2016}, BRENDA~\cite{Placzek2017} and
QSAR DataBank~\cite{Ruusmann2015} databases.
A number of such tools were initially developed using older versions
of the CDK and are updated to new releases as they are made
available. Examples include Bioclipse~\cite{spjuth2007bioclipse,
spjuth2009bioclipse} and
AMBIT~\cite{jeliazkova2011ambit,jeliazkova2011ambitsmarts,kochev2013ambit}. The
CDK has also played a role in a number of chemical studies, such as finding
the maximally bridging rings in chemical structures~\cite{Marth2015},
prediction of organic reactions~\cite{Segler2016}, and bioactivities
of compounds~\cite{Alvarsson2016}.

>>>>

The open source cheminformatics community has made significant steps
forward recently~\cite{OBoyle2011b} as evidenced by the growing number
of tools and underlying toolkits, along with the usage of these
software components in a variety of applications.
The Chemistry Development Kit (CDK) is one of the tools developed
under the aegis of the Blue Obelisk, a movement promoting Open Data,
Open Source, and Open Standards in chemistry~\cite{OBoyle2011b,Guha2006}.
Previously documented CDK versions have been widely adopted~\cite{Steinbeck2003,Steinbeck2006}.
Use of the CDK ranges from inclusion of CDK functionality in
wrapper platforms such as Cinfony~\cite{OBoyle2008}, incorporation
within the R environment (rcdk~\cite{Guha2007}), and as plugins for
Taverna~\cite{Truszkowski2011}, 
KNIME~\cite{Beisken2013}, Cytoscape (ChemViz2~\cite{ChemViz2}), and for
Microsoft Excel (LICSS~\cite{Lawson2012}).
In contrast to scenarios that have made CDK functionality available in
larger systems, a number of projects have employed the CDK as a
general cheminformatics toolkit. Examples include 
jCompoundMapper~\cite{Hinselmann2011}, ScaffoldHunter~\cite{wetzel2009interactive,Klein2013}, OMG~\cite{Peironcely2012},
PaDEL~\cite{yap2011padel}, ChemDes~\cite{Dong2015},
ReactPRED~\cite{ReactPRED}, SMSD~\cite{Rahman2009,Rahman2014,Rahman2016},
WhichCyp~\cite{Rostkowski2013}, MetaPrint2D~\cite{Carlsson2010}, MetFrag~\cite{Wolf2010},
and the IUPHAR/BPS Guide to Pharmacology~\cite{Southan2016}, BRENDA~\cite{Placzek2017} and
QSAR DataBank~\cite{Ruusmann2015} databases.
A number of such tools were initially developed using older versions
of the CDK and are updated to new releases as they are made
available. Examples include Bioclipse~\cite{spjuth2007bioclipse,
spjuth2009bioclipse} and
AMBIT~\cite{jeliazkova2011ambit,jeliazkova2011ambitsmarts,kochev2013ambit}. The
CDK has also played a role in a number of chemical studies, such as finding
the maximally bridging rings in chemical structures~\cite{Marth2015},
prediction of organic reactions~\cite{Segler2016}, and bioactivities
of compounds~\cite{Alvarsson2016}.
